---
title: "P5 Submission"
author: "Sergio Comuñas"
date: "30 de agosto de 2017"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(knitr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:/OneDrive/Perfil/Documentos/Proyectos/2017-08 Data Analyst Nanodegree (udacity)/P5 Exploratory Data Analysis/Project")
redwine <- read.csv("WineQualityRed.csv", stringsAsFactors = FALSE)
```

I've chosen Red Wine Quality dataset because here in Spain there's a lot of good
wines and it's interesting for me.

In the "WineQualityRedDescription.txt" file you can see the dataset description.

# Univariate Plots Section

We can observe the dimensions of the dataset:

```{r echo=FALSE, Dimensions}
cat(length(redwine$X), "rows and", length(redwine), "features")
```

Also we can view a short summary in terms of each feature and the numbering
values of it:

```{r echo=FALSE, Summary}
str(redwine)
summary(redwine)
```

The first feature I'm interested in is "Quality" that gives us the wine score
between 0 and 10):

```{r echo=FALSE, HistQuality}
qplot(redwine$quality, geom="histogram", binwidth = 1) +
ggtitle("Quality Feature Distribution") +
xlab("Quality (1-10)") +
ylab("Number of instances")
```

The next one gives us a vision about alcohol density:

```{r echo=FALSE, HistAlcohol}
qplot(redwine$alcohol, geom="histogram", binwidth = 0.1) +
ggtitle("Alcohol Feature Distribution") +
xlab("Alcohol (%)") +
ylab("Number of instances")
```

Also I want to take a look to the sulphates feature because here in my house
always have discussion about the quantity of each one:

```{r echo=FALSE, HistSulphates}
qplot(redwine$sulphates, geom="histogram", binwidth = 0.03) +
ggtitle("Sulphates Feature Distribution") +
xlab("Sulphates (%)") +
ylab("Number of instances")
```

# Univariate Analysis

This dataset is structured as a punctuation table for a list of about 1200 wines
anb his relationed features.

For me the main feature is quality. In most of the cases quality must converge
in a combination of other features to give most feeling to the wine.

Other variables in this dataset are the ones reliable to reality. The
ones that every people can feel driking the wine. Then: Acidity, Residual Sugar,
Chlorides, Sulfur Dioxide, Density, pH, Sulphates and Alcohol.

I've simplified the dataset creating total.acidity (sum of the three features)
and removing free Sulfur Dioxide feature (taking in consideration only Total SD 
feature).

Again there's a median detail about each feature:

```{r echo=FALSE, NewFeatures}
redwine$acidity = redwine$fixed.acidity + redwine$volatile.acidity + 
    redwine$citric.acid
redwineClean = redwine[, c("X", "quality", "acidity", "residual.sugar", 
                           "chlorides", "total.sulfur.dioxide", "density", 
                           "pH", "sulphates", "alcohol")]

summary(redwineClean)
```


# Bivariate Plots Section

First of all I want to see relation between quality and the most important 
features.

Here we can see relation between quality and alcohol:

```{r echo=FALSE, PlotAlcohol}
ggplot(aes(quality, alcohol, group = quality), data = redwine) + geom_boxplot()+
ggtitle("Quality vs Alcohol") +
xlab("Quality") +
ylab("Alcohol (%)")
```

Another plot we can observe is quality vs acidity because in red wines acidity
is relationed with bad ones:

```{r echo=FALSE, PlotAcidity}
ggplot(aes(quality, acidity, group = quality), data = redwine) + geom_boxplot()+
ggtitle("Quality vs Acidity") +
xlab("Quality") +
ylab("Acidity")
```

The last relation I want to observe is Sugar vs alcohol. I've heard from child
that sugar is the ones come to alcohol after fermentation. Then, mainly the 
less sugar remaining would become more alcohol. Let's see:

```{r echo=FALSE, PlotSugar}
ggplot(aes(residual.sugar, alcohol), data = redwine) + geom_point() +
ggtitle("Residual Sugar vs Alcohol") +
xlab("Residual Sugar (grams/liter)") +
ylab("Alcohol (%)")
```


# Bivariate Analysis

We can do a first approach of the mean of each feature for every punctuation in
quality:

```{r echo=FALSE, TableDatabyQuality}
DataByQuality = aggregate(redwineClean[, 2:10], by = list(redwineClean$quality), 
                          FUN = "mean")[, 2:10]
kable(DataByQuality, digits = 3, col.names = c("Quality", "Acidity", 
                                               "Residual Sugar", "Chlorides", 
                                               "Sulfur Dioxide", "Density", 
                                               "pH", "Sulphates", "Alcohol"))
```

As you can see:
- There's a strong relation between alcohol and quality but not as I expected:
for more alcohol more quality (in approx)
- There's no relation between acidity and quality
- It's not true about less sugar more alcohol. Maybe there's rain types with
more sugar in origin??


# Multivariate Plots Section


> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

First of all come again with the strong relation we have: quality vs alcohol. In
this case we add sulphates feature in order to see if increases the correlation.

Here you can see the plot:

```{r echo=FALSE, PlotAlcoholSulphates}
ggplot(aes(quality, alcohol, alpha = sulphates), data = redwine) + geom_point()+
ggtitle("Residual Sugar vs Alcohol vs Sulphates") +
xlab("Residual Sugar (grams/liter)") +
ylab("Alcohol (%)") +
labs(alpha = "Sulphates")
```

You can see there's no apparent relation for sulphates. Only can observe there's
extreme low sulphates rate on better and on poor wines. I suspect high quality
wines there's no sulphates because of they think on better quality rates and in
lower quality wines because they want not to spent so much in order to have
affordable bottles.

We can do something similar but taking in density feature instead sulphates:

```{r echo=FALSE, PlotAlcoholDensity}
ggplot(aes(quality, alcohol, alpha = density), data = redwine) + geom_point() +
ggtitle("Residual Sugar vs Alcohol vs Density") +
xlab("Residual Sugar (grams/liter)") +
ylab("Alcohol (%)") +
labs(alpha = "Density")
```

You can see there's a relation between density and quality. All of the High
Class Wines have low density rate. There is not so strong relation as alcohol
but thinking on a future prediction model (for example) is sure it's a feature
to include.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
qplot(redwine$quality, geom="histogram", binwidth = 1) +
ggtitle("Quality Feature Distribution") +
xlab("Quality (1-10)") +
ylab("Number of instances")
```

### Description One

First one is the Quality distribution plot. For me is so interesting in terms
that have a near normal distribution. You can see there's a lot mid-market wines
and a poor high class wines. Also you can see there's a little growing on the 
right. I think it must be related to there's so few bad wines but also few great
ones.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(residual.sugar, alcohol), data = redwine) + geom_point() +
ggtitle("Residual Sugar vs Alcohol") +
xlab("Residual Sugar (grams/liter)") +
ylab("Alcohol (%)")
```

### Description Two

This one is so special for me because it crashes a family theory I've heard for
years. The more sweet wine the more alcohol you have. It's clearly not true
because no clear relation between the two features.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(quality, alcohol, alpha = density), data = redwine) + geom_point() +
ggtitle("Residual Sugar vs Alcohol vs Density") +
xlab("Residual Sugar (grams/liter)") +
ylab("Alcohol (%)") +
labs(alpha = "Density")
```

### Description Three

I've selected this one because clearly shows us the relation between alcohol and
quality but also shows us that there is a lot of range for mid market wines.
There's a strong relation but not enough to ensure high alcohol wines are
better.

Also you can see there's a significan low density wines for High Class ones.
This may be useful for a stronger analysis.

------

# Reflection

There's a lot of wine classes, this is not new. The most of them are mid-market
ones. It's not clear for have an strong model to predict if a wine will be good
or not. Here in Spain the wine people thell us that it's not science, it's a 
culture process. I'm not sure about that but with that data can conclude it
would not be easy to define the relation.

For me the next steps would be investigate the correlation for quality vs the 
rest of the features. Maybe would star with a ggpairs plot or something similar
to include more mathematics and try to model the strong relations.

In the top of my thougs would be great to have a model that can predict quality
for the wine in terms of prediction. Most of wine people can predict: acidity, 
sulphates entered, etc. They will be crazy if can predict income for every year
and Marketing actions to be done...